Co-methylation

### 2015-02-18 Rmarkdown/comethylation.Rmd
A few TODOs

@@

These plots appear to show that there is a dramatic increase in within-fragment co-methylation at precisely these $IPD$s for paired-end samples from the Lister dataset. However, I am very skeptical as to the biological relevance of these results.

@@

As discussed in Section \ref{sec:m-bias}, the incomplete removal of M-bias is a real problem for reads that have undergone pre-trimming because this destroys the relationship between read-position and sequencing cycle. __FIGURE__ shows that there is considerable variation in the read lengths of the _ADS_ sample, indicative of aggressive pre-trimming of these reads (__TODO: Histogram of read lengths__).

To investigate my suspicion that these results are due to the incomplete removal of M-bias, I used an aggressive filtering strategy that attempts to only use read-positions that are unlikely to suffer from M-bias. I applied this to the _ADS_ sample. In this analysis I ignored read-positions $1-20$ and $55-80$ of both _read 1_ and _read 2_. Unfortunately, this is a situation where the M-bias filtering strategy of `methtuple` is a disadvantage. `methtuple` is designed to filter out specific read-positions and cannot 'exclude the first/last $n$ bases'. For example, using this aggressive strategy will have a detrimental effect for a read that has been pre-trimmed to $30$ bp, of which the last $10$ bp are affected by M-bias, because this filtering strategy will __retain__ read-positions $21-30$, which are exactly those affected by M-bias. Nevertheless, it is the best that can be done with the current version of the software[^methtuple_refactoring] and with such aggressively pre-trimmed data.

@@

The results for the aggressively trimmed _ADS_ sample are shown in __FIGURE__.

__TODO: Comment on these results and how we should interpret the Lister co-methylation results (particularly NIL >= 0)__

__TODO: Summarise the M-bias issue__

>>>>

# TODOs

- Try out Matthew Stephen's adaptive hierarchical shrinkage method on the estimated log odds ratios (see [https://github.com/stephens999/ash](https://github.com/stephens999/ash)).

@@

__TODO: I need to re-write this section (bold). I'm really not yet sure if this result is biology or an artefact. My instinct says artefact but I haven't yet been able to prove this either way to my satisfaction
These plots appear to show that there is a dramatic increase in within-fragment co-methylation at precisely these $IPD$s for paired-end samples from the Lister dataset. However, I am very skeptical as to the biological relevance of these results.

@@

As discussed in Section \ref{sec:m-bias}, the incomplete removal of M-bias is a real problem for reads that have undergone pre-trimming because this destroys the relationship between read-position and sequencing cycle. FIGURE shows that while there is considerable variation in the read lengths of the _ADS_ sample, which is indicative of aggressive pre-trimming of these reads, most reads are full length or close to it (TODO: Histogram of read lengths).

To investigate my suspicion that these results are due to the incomplete removal of M-bias, I used an aggressive filtering strategy that attempts to only use read-positions that are unlikely to suffer from M-bias. I applied this to the _ADS_ sample. In this analysis I ignored read-positions $1-20$ and $55-80$ of both _read 1_ and _read 2_. Unfortunately, this is a situation where the M-bias filtering strategy of `methtuple` is a disadvantage. `methtuple` is designed to filter out specific read-positions and cannot 'exclude the first/last $n$ bases'. For example, using this aggressive strategy will have a detrimental effect for a read that has been pre-trimmed to $30$ bp, of which the last $10$ bp are affected by M-bias, because this filtering strategy will _retain_ read-positions $21-30$, which are exactly those affected by M-bias. Nevertheless, it is the best that can be done with the current version of the software[^methtuple_refactoring] and with such aggressively pre-trimmed data.

@@

The results for the aggressively trimmed _ADS_ sample are shown in FIGURE.

- TODO: Comment on these results and how we should interpret the Lister co-methylation results (particularly NIL >= 0)
- TODO: Summarise the M-bias issue
__
